\LARGE \bf Informed sampling by Markov Chain Monte Carlo methods

### 2017-02-24 paper.tex
added appendix

\appendix
\section{HMC}
\os{polish}

In consider of efficiency, Hamiltonian Monte Carlo(HMC)~\cite{N11} is thus introduced to replace random walk with Hamiltonian dynamics, which means that future states in the Markov chain is proposed from simulating system dynamics instead of sampling from a jumping distribution.
It is named ``Hamiltonian, because the concept origins from Hamiltonian dynamics, which defines how system dynamics is determined by system energy.
Define the system dynamics of an object as the position $ x $ and the momentum $ q $ over time $ t $.
We have a system energy $ E $ 
%\os{change to E}
of the object that consists of kinetic energy $ K $ and potential energy $ U $, which is written as $ E(x,q) = U(x) + K(q) $.
The kinetic energy $ K(q) $ is determined by the momentum $ q $ of the object, while the potential energy $ U(x) $ is determined by the position $ x $ of the object.
The Hamiltonian dynamics is determined by the gradients of the system energy, which is
$ \frac{dx}{dt} = \frac{\partial E}{\partial q} $ and $ \frac{dq}{dt} = - \frac{\partial E}{\partial x} $.
%
The canonical distribution is often choose to bridge a system energy with a probability distribution, which is written as $ P(x, q) = \frac{1}{Z} \exp ( -E(x,q) / T ) $.
If two different states have the same energy, the probabilities of them are equivalent.
It has been proved that properties of Hamiltonian dynamics ``leaves the canonical distribution invariant"~\cite{N11}, which means $ P(x, q) $ equals to a stationary distribution in MCMC.
Thus, we can generate samples $ [ x , q ]^T $ from joint distribution $ p(x, q) $.
%
Given a target distribution $ \pi(x) $, the potential energy is defined in the form $  U(x) = - \log ( \pi(x) T ) $.
We can have $ P(x, q) = \pi(x) \frac{1}{Z'} \exp ( -K(q) / T ) $, which indicates that $ pi(x) $ is a marginal distribution of $ P(x, q) $.
In HMC, we sample $ [ x ,q ] $ from $ P(x, q) $ and discard the momentum $ q $ so that the set of samples $ x $ equals to sampling from the marginal distribution $ \pi(x) $.
The momentum $ q $ is used as an auxiliary variable.
Avoiding random walk helps samplings converge to the desired distribution $ \pi(x) $ in higher rate~\cite{N11}.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Old text that may be re-used
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ignore{
In order to take advantage of these powerful methods, we propose to reformulate the cost surface as a probability density. Specifically, we consider a Boltzmann distribution in order to formulate the cost function using flexible energy based methods. Because Monte Carlo methods only consider the ratio of the probabilities, we do not need to consider the problem of estimating the intractable partition function.

>>>>

\appendix
\section{HMC}
In this section we provide background on Hamiltonian Monte Carlo (HMC). 
For a comprehensive description, see~\cite{N11},
for intuition, see~\cite{ZRDPKDBS13}.
%Some of the notation in this section

MCMC algorithms typically make use of random-walks to convergence to a target stationary distribution.
This is often inefficient, resulting in slow mixing\footnote{The \emph{mixing time} of a Markov chain is the time until it is ``close'' to its steady state distribution.}.
HMC is an MCMC method that adopts physical-system dynamics (specifically Hamiltonian dynamics) rather than a probability distribution to produce the next state in the Markov chain.

\subsection{Hamiltonian dynamics}
Hamiltonian dynamics describe an object's motion in terms of its position $x$ and momentum $q$ at some time $t$. 
Each position is associated with a potential energy $U(x)$.
Similarly, each momentum is associated with a kinetic energy $K(q)$.
The total energy (also called the Hamiltonian) of the system is constant and is defined as
$$
E(x, p) = U(x) + K(q).
$$
Hamiltonian dynamics describe how the kinetic and potential energies are converted one to the other as the object moves throughout a system in time. 
This is done via a set of differential equations known as the Hamiltonian equations:
$$
\frac{dx}{dt} = \frac{\partial E}{\partial q}, 
\hspace{5mm}
\frac{dq}{dt} = - \frac{\partial E}{\partial x}. $$
%
%The Hamiltonian equations describe an object’s motion in time, which is a continuous variable. In order to simulate Hamiltonian dynamics numerically, the Leapfrog method is typically used.

\subsection{From Hamiltonian dynamics to probability distribution}
To sample from a desired distribution~$P$ using HMC, we need to convert the distribution to an energy function and vice-versa. This is done using the concept of a canonical distribution from statistical mechanics~\cite{N11}. Given some energy function,~$E$, the canonical distribution over states has the Boltzman probability distribution $P = \frac{1}{Z}\exp (-E/T)$, where $Z$ is a normalizing constant and~$T$ is the temperature of the system.
Notice that in our setting
$$
P(x,q) 
\propto e^{-E(x,q)} 
= e^{-U(x)} \cdot e^{-K(q)}
\propto P(x) \cdot P(q).
$$
This means that the distributions $P(x)$ and $P(q)$ are independent. Therefore, we can use Hamiltonian dynamics to sample from the joint distribution over $q$ and $x$ and ignore the momentum contributions.
Because the two distributions are independent, we can introducing auxiliary variables to represent the momentum and choose any distribution from which to sample the momentum variables. A common choice is to use a zero-mean Normal distribution with unit variance.

\subsection{Hamiltonian Monte Carlo}
HMC uses Hamiltonian dynamics to explore the distribution $P(x)$ defined by $U(x)$. 
Starting at an initial state $(x_0, q_0)$, 
(here, $q_0$ is the axillary variable used to simulate the momentum), Hamiltonian dynamics are simulated for a short time using the Leap Frog method. 
Let $(x^*, q^*)$ be the state at the end of the simulation. 
This state is accepted using an update rule analogous to the Metropolis acceptance criterion~\cite{CG95}. 
Specifically, if the probability of $(x^*, q^*)$ after Hamiltonian dynamics is greater than probability of the state prior to the Hamiltonian dynamics, then $(x^*, q^*)$ is accepted.
Otherwise, $(x^*, q^*)$ is accepted randomly. 
If $(x^*, q^*)$ is rejected, the next state of the Markov chain is set as the state at $(t-1)$.

To explore the entire space, the dynamics are randomly perturbed. This is done by drawing a random momentum from $p(q)$ before running the dynamics prior to each sampling iteration $t$. Combining these steps, sampling random momentum, followed by Hamiltonian dynamics and Metropolis acceptance criterion defines the HMC algorithm.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Old text that may be re-used
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ignore{
In order to take advantage of these powerful methods, we propose to reformulate the cost surface as a probability density. Specifically, we consider a Boltzmann distribution in order to formulate the cost function using flexible energy based methods. Because Monte Carlo methods only consider the ratio of the probabilities, we do not need to consider the problem of estimating the intractable partition function.
